---
title: "Portfolio"
execute: 
  dir: r-calbow-portfolio
---

About this site

Hi! My name is Riley Calbow. I am a fourth year student at California State University Monterey Bay studying marine science and statistics. I participate in undergraduate research on intertidal ecosystems and phycological changes over time. My expected graduation date is fall 2026 and I plan to contribute to the conservation of coastal ecosystems in the near future.

### **Data Description**

1.  Identify your data source.

The Haupt Lab and MARINe/PISCO dataset

2.  Describe your data, including variables and data types.

List of variables in the dataframe:

Site: character

Date: numeric

Transect #: numeric

Latitude and Longitude for start and end transect: numeric

Quadrat #: numeric

Sea or Land Side: character

One column for the percent cover each organism observed: includes various functional groups of algae, kelp, mussels, anemones, barnacles, and other: numeric

Urchin density: numeric

Who entered the Data: character

Notes: character

These datasets include the observed percent cover of various organisms within intertidal sites across the Monterey Peninsula

3.  Identify the research questions you want to answer.

How have both urchin density and algae cover changed over time and how do they appear to have interacted with each other? Has one increased and the other decreased, or vice versa?

### **Data Visualization**

1.  What do you want your final visualizations to look like?

I definitely want to include boxplots of each site to show urchin density. This is part of my preliminary data visualization and analysis for my capstone proposal so ideally I would like to include an NMDs plot which I have some experience with.

2.  What do you want to highlight in your final visualizations in order to answer your research questions? How do you plan to do that?

I’d like to highlight change over time and how different sites have changed. I’d like to just get a general idea of how intertidal ecosystems have changed over time and maybe highlight specific sites that saw large changes. Maybe using colors or ordering my visualizations in a way that draws more attention to them.

3.  What is missing from your data or would need to change in your data to create these visualizations?

I mainly think I will need to join datasets together to create comprehensive visualizations as opposed to several smaller ones.

### **Data Cleaning**

The answer to at least three of these questions should be “YES” for the data to meet the necessary standards to demonstrate your cleaning. Your data source should not be an already perfectly prepared data set.

1.  Do you need to reformat any variables into different types (e.g. factors, time, dates, strings)? Or remove information from variable values?

No

2.  Do you need to deal with any missing data, especially missing data coded other than NA?

No

3.  Do you need to filter your data? How?

Yes, by sites or location and for specific species

4.  Do you need to create any new variables? What variables? How?

Yes, a total percent cover may be helpful for analysis as well as a proportion of percent cover instead of just percent cover

5.  Do you need to add new data (join) to your data? What data? How?

Yes, I have several csvs that will need to be joined together so I can have urchin and algae data in the same csv.

6.  Do you need pivot your data in any way? Why? How?

No

7.  Do you need to summarize any of the variables? Which ones? How?

Yes, urchin and algae percent cover will likely be summarized as a mean proportion for each year we have data

8.  What other aspects of your data need to be “fixed” in order to make your data visualizations?

Last time I used this data many of the site names had different letters capitalized, different abbreviations, etc so names will need to be cleaned.

Most will answer yes to the following for making your programming more efficient using select(), but you should have three other yeses above.

9.  Are there any variables you can exclude from your data?

Yes

# Data Cleaning

```{r setup}
library(tidyverse)
library(lubridate)
library(tidyr)
library(dplyr)

```

```{r read in data}
urchin <- read.csv("UrchinDensity.csv")
algae <- read.csv("AlgalCover_2021.csv")
```

```{r check density data}
str(urchin)
unique(urchin$Site) #check if we have duplicates of sites spelled differently/spelled wrong etc.
urchin_clean <- urchin |>
  mutate(Site = replace(Site,Site == "Pecadero Point", "Pescadero Point")) |>
  mutate(Site = replace(Site,Site == "Hopkins ", "Hopkins")) |>
  mutate(Site = replace(Site,Site == "China Rock ", "China Rock")) |> 
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) #change names that are spelled wrong or have an extra space so they are in the same site name
#remove columns I don't need for analysis
unique(urchin_clean$Site) #make sure it worked
```

```{r}
unique(algae$Site)
algae_clean <- algae |>
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) #remove columns I won't need to include in analysis
```

```{r put in clean data folder}
#write_csv(urchin, file = "Algae-Project/UROC-Spring-2026-Data/clean-data/urchin_clean.csv")
#clean_urchin <- read_csv("Algae-Project/UROC-Spring-2026-Data/clean-data/urchin_clean.csv")

#write_csv(algae, file = "Algae-Project/UROC-Spring-2026-Data/clean-data/algae_clean.csv")
#clean_algae <- read_csv("Algae-Project/UROC-Spring-2026-Data/clean-data/algae_clean.csv")
```

```{r join data}
#clean_urchin |> 
 # full_join(clean_algae, by = "Site") #joining the two dataframes for now
#may change how they are joined/clean certain columns once I start making visualizations depending on what works best

```
