---
title: "Project: Example for Class"
format: html
execute: 
  echo: false
  warning: false
  message: false
embed-resources: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(lubridate)
library(tidyr)
library(dplyr)
library(paletteer)
```

## Data Description

This data was collected by the CSUMB Rocky Intertidal Lab in 2021 including the percent cover of various algae types and the density of purple sea urchins across 7 sites; Pescadero Point, Carmel Point, Hopkins, China Rock, Sea Palm, Cannery Row. Algae percent cover was estimated by using one meter quadrats and identifying primary algae functional groups and sessile organisms.

## Data Cleaning

To clean these data I first fixed any spelling errors in sites that caused the same sites to show up multiple times with different spelling by combining them into one label for each site. I then removed all columns I did not need for my analysis. To ensure I could include all algae types in a boxplot, I pivoted the all algae type columns longer into one column for which algae type a percent cover estimate is for called "algae_type" and created a new column for the percent cover estimates called "perc_cover". Finally, to make sure the algae boxplot showed the full functional group names, not the abbreviations, I changed the abbreviated names of algae groups to the full name.

```{r}
#| label: data-cleaning
#| include: false
#this shoud be hidden from view and no seen in the website it should include your data cleaning and you should export the clean data
urchin <- read.csv("UrchinDensity.csv")
algae <- read.csv("AlgalCover_2021.csv")

```

```{r check density data}
#| include: false
str(urchin)
unique(urchin$Site) #check if we have duplicates of sites spelled differently/spelled wrong etc.
urchin_clean <- urchin |>
  mutate(Site = replace(Site,Site == "Pecadero Point", "Pescadero Point")) |>
  mutate(Site = replace(Site,Site == "Hopkins ", "Hopkins")) |>
  mutate(Site = replace(Site,Site == "China Rock ", "China Rock")) |> 
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) #change names that are spelled wrong or have an extra space so they are in the same site name
#remove columns I don't need for analysis
unique(urchin_clean$Site) #make sure it worked
```

```{r}
#| include: false
unique(algae$Site)
algae_clean <- algae |>
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) |>
  pivot_longer( cols = 10:19,
    names_to = "algae_type",
               values_to = "perc_cover") |>
  mutate(algae_type = fct_recode(factor(algae_type),
                               "Articualted Coralline" = "AC",
                               "Crustose Coralline" = "Cr.Cor",
                               "Crustose Non-Coralline" = "Cr.NC",
                               "Filamentous" = "Fila.",
                               "Corticated" = "Cort.",
                               "Leathery" = "Leath.",
                               "Seagrass" = "SG",
                               "Bare Rock" = "BR"))#remove columns I won't need to include in analysis
#Rename abbreviated variables to actual algae functional group names
unique(algae_clean$Site) 
```

```{r}
#| include: false
#write in clean data into the folder
write_csv(urchin_clean, file = "clean-data/urchin_final.csv")
urchin_final <- read_csv("clean-data/urchin_final.csv")

write_csv(algae_clean, file = "clean-data/algae_final.csv")
algae_final <- read_csv("clean-data/algae_final.csv")
```

## Urchin Density

This visualization shows boxplots of the urchin density of seven different sites across the Monterey Bay ordered from lowest urchin density to highest urchin density.Pescadero Point shows the highest urchin density at a median of around 180 individuals recorded, Cannery Row and Carmel Point with 50 and 40 individuals respectively. All other sites had an average number of urchins close to 0.
Here is a description of the first visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-1
#| fig-cap: "Boxplots showing the urchin density in individuals at each of the sites monitored. Pescadero Point had the highest number of urchins around a 180 median."
#| fig-alt: "Boxplots shoing the urchin density in individuals on the y axis of seven sites across Monterey Bay on the x axis. They are in order from left to right from lowest to highest number of urchins observed. Pescadero Point has the highest median and range with a median around 180 and a range form 350 to 0, with two outliers around 600. Cannery Row has the next highest median around 50 with a range from 280 to  with one outlier at 300. Carmel point is the third highest with a median around 40 and a range from 100 to 0, and 4 outliers at 300, 400, and roughly 450. All other sites including Sea Palm, Point Pinos, China Rock, and Hopkins all have a median close to 0 and a range of approximately 50 to 0 or smaller, with outliers at the highest around 180. "
#| fig-width: 6
#| fig-height: 5
#| fig-align: center

#create plot of iris dataset
ggplot(urchin_final, aes(x = fct_reorder(Site, S_purpuratus), 
                         y = S_purpuratus, 
                         fill = as.factor(Site))) +
  geom_boxplot() +
  labs(title = "Urchin Density Across Sites", 
       x = "Site", 
       y = "Urchin Density (# indv.)") +
 scale_fill_paletteer_d("nationalparkcolors::Arches")+
  theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = .75),
      legend.position = "none")
```

## Algae Percent Cover in Monterey

This visualization shows the percent cover of each algae functional group across intertidal zones of Monterey Bay. Similar to the first one it shows the percent cover of each group for all seven of the sites. Each site show a different combination of abundance of various algae groups and some are dominated by different functional groups than others or even by bare rock. Hopkins, China Rock, and Carmel Point are dominated by leathery algae. Cannery Row and Pescadero Point are dominated by bare rock, meaning there is a lack of algae at those sites. Point Pinos and Sea Palm are dominated by Crustose Non-Corraline, then bare rock and leathery algae respectively.


```{r}
#| label: viz-2
#| fig-cap: "This is a caption for the second visualization"
#| fig-alt: "This is an alternative text for the second visualization"
#| fig-width: 8
#| fig-height: 6
#| fig-align: center

#create plot of cars data set
ggplot(algae_final, aes(x = Site, y = perc_cover, fill = algae_type)) +
  geom_boxplot() +
  labs(title = "Algae Percent Cover by Site", 
       x = "Site", 
       y = "Percent Cover (/100%)",
       fill = "Algae Group") +
    scale_fill_discrete(labels = c("AC" = "Articulated Coralline"))+
  scale_fill_paletteer_d("ggthemes::Classic_Cyclic")+
  theme_minimal()
```
