---
title: "Project: Example for Class"
format: html
execute: 
  echo: false
  warning: false
  message: false
embed-resources: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(lubridate)
library(tidyr)
library(dplyr)
library(paletteer)
```


## Data Description

This data was collected by the CSUMB Rocky Intertidal Lab in 2021 including the percent cover of various algae types and the density of purple sea urchins across

## Data Cleaning

Here I describe in plain speak what I did to clean the data. Do not show code or use technical terms. I want to be able to read this and understand what I did.

```{r}
#| label: data-cleaning
#this shoud be hidden from view and no seen in the website it shouldl include your data cleaning and you should export the clean data
urchin <- read.csv("UrchinDensity.csv")
algae <- read.csv("AlgalCover_2021.csv")

```

```{r check density data}
str(urchin)
unique(urchin$Site) #check if we have duplicates of sites spelled differently/spelled wrong etc.
urchin_clean <- urchin |>
  mutate(Site = replace(Site,Site == "Pecadero Point", "Pescadero Point")) |>
  mutate(Site = replace(Site,Site == "Hopkins ", "Hopkins")) |>
  mutate(Site = replace(Site,Site == "China Rock ", "China Rock")) |> 
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) #change names that are spelled wrong or have an extra space so they are in the same site name
#remove columns I don't need for analysis
unique(urchin_clean$Site) #make sure it worked
```

```{r}
unique(algae$Site)
algae_clean <- algae |>
  select(-c("Investigators", "Entered.By", "Checked.By", "Notes")) |>
  pivot_longer( cols = 10:19,
    names_to = "algae_type",
               values_to = "perc_cover") |>
  mutate(algae_type = fct_recode(factor(algae_type),
                               "Articualted Coralline" = "AC",
                               "Crustose Coralline" = "CR.Cor",
                               "Crustose Non-Coralline" = "8"))#remove columns I won't need to include in analysis
#FINISH RENAMING THE VARIABLES THIS IS WHAT YOU WERE WORKING ON

```




```{r}
#write in clean data into the folder
write_csv(urchin_clean, file = "clean-data/urchin_final.csv")
urchin_final <- read_csv("clean-data/urchin_final.csv")

write_csv(algae_clean, file = "clean-data/algae_final.csv")
algae_final <- read_csv("clean-data/algae_final.csv")
```



## Visualzation 1

Here is a description of the first visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-1
#| fig-cap: "This is a caption for the first visualization"
#| fig-alt: "This is an alternative text for the first visualization"
#| fig-width: 6
#| fig-height: 5
#| fig-align: center

#create plot of iris dataset
ggplot(urchin_final, aes(x = fct_reorder(Site, S_purpuratus), 
                         y = S_purpuratus, 
                         fill = as.factor(Site))) +
  geom_boxplot() +
  labs(title = "Urchin Density Across Sites", 
       x = "Site", 
       y = "Urchin Density (m^2)") +
 scale_fill_paletteer_d("nationalparkcolors::Arches")+
  theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = .75),
      legend.position = "none")
```

## Visualization 2

Here is a description of the second visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-2
#| fig-cap: "This is a caption for the second visualization"
#| fig-alt: "This is an alternative text for the second visualization"
#| fig-width: 8
#| fig-height: 6
#| fig-align: center

#create plot of cars data set
ggplot(algae_final, aes(x = Site, y = perc_cover, fill = algae_type)) +
  geom_boxplot() +
  labs(title = "Algae Percent Cover by Site", 
       x = "Site", 
       y = "Percent Cover (/100%)",
       fill = "Algae Group") +
    scale_fill_discrete(labels = c("AC" = "Articulated Coralline"))+
  scale_fill_paletteer_d("ggthemes::Classic_Cyclic")+
  theme_minimal()
```